On the number of tiles visited by a\\ line segment on a rectangular grid

### 2023-07-01 paper.tex
Errata corrected, for second revision

For any $i, j \geq 2$ with $i+j-1=\tiles$, the segment has a discrete bounding rectangle with dimensions $i, j$ if and only if its second endpoint is in the tile with lower-left corner $(i-1,j-1)$ or in one of the other three symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$, i.e.~those with lower-left corners $(i-1,1-j)$, $(1-i,1-j)$ and $(1-i,1-j)$. An example is shown in Figure~\ref{fig: probmax_symmetric_tiles}. Consequently, $\probmax_{i,j}(\len)$ can be obtained by computing the probability that the second endpoint is in the tile with lower-left corner $(i-1,j-1)$ and multiplying by $4$.

@@

By Proposition~\ref{prop: len ineq i j}, the tile with lower-left corner $(i-1,j-1)$ can be visited only if
\begin{equation}
\label{eq: i-2 j-2 len}
(i-2)^2+(j-2)<\len^2.
\end{equation}
Thus $\probmax_{i,j}(\len)=0$ if this restriction does not hold. In the following it will be assumed that it does.

@@

Using the coordinate axes defined in Figure~\ref{fig: probmax_asint_turned}, the circular arc satisfies the equation $y = \funcirc(x)$ with
\begin{equation}
\label{eq: funcirc}
\funcirc(x) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-x^2},
\end{equation}
where the arc height $h_\tiles$ is 
\begin{equation}
\label{eq: h tiles}
h_\tiles = \len_\tiles - \frac{\tiles-3}{\sqrt{2}}.
\end{equation}
The $y$ coordinate of the arc for $x=x_{\tiles,n}$ will be denoted as $y_{\tiles,n} = \funcirc(x_{\tiles,n})$. The half-width of the arc, $w_\tiles$, is computed in the proof of Theorem~\ref{theo: probmax, sq} (or from the right triangle in Figure~\ref{fig: probmax_asint_turned}), and using \eqref{eq: len tiles} it can be expressed as
\begin{equation}
\label{eq: w}
w_\tiles = \sqrt{\len_\tiles^2-\frac{(\tiles-3)^2}{2}} =
\begin{cases}
\displaystyle
\sqrt{\tiles-2} & \text{ for $\tiles$ odd} \\
\displaystyle
\sqrt{\tiles-5/2} & \text{ for $\tiles$ even}.
\end{cases}
\end{equation}
The inequality
\begin{equation}
\label{eq: sqrt asympt}
\sqrt{1+\genvar} < 1 + \frac \genvar 2 \quad\text{for } s \in \mathbb R, s \neq 0
\end{equation}
implies that
\begin{equation}
\label{eq: asympt w}
w_\tiles \leq \sqrt{\tiles-2} < \sqrt{\tiles} - 1/\sqrt\tiles.
\end{equation}
Similarly, \eqref{eq: len tiles} and \eqref{eq: sqrt asympt} yield, considering that $\tiles \geq 3$,
\begin{equation}
\label{eq: asympt len tiles}
\frac{\tiles-2}{\sqrt{2}} \leq \len_\tiles < \frac{\tiles-2}{\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}
\leq \frac{\tiles-3/2}{\sqrt{2}}.
\end{equation}
The arc height $h_\tiles$ is bounded from \eqref{eq: h tiles} and \eqref{eq: asympt len tiles} as
\begin{equation}
\label{eq: h bound}
\frac {\sqrt{2}} 2 \leq h_\tiles <
% \frac 1 {\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)} \leq
\frac{3\sqrt{2}}{4}.
\end{equation}
The number $N_\tiles$ of reachable tiles shown in the figure (circular markers) is easily obtained from $w_\tiles$ as
\begin{equation}
\label{eq: N tiles}
N_\tiles = 
\begin{cases}
\displaystyle
2\left\lceil \frac{w_\tiles}{\sqrt{2}} \right\rceil - 1 = 2\left\lceil \sqrt{\frac{\tiles-2}{2}} \right\rceil - 1 & \text{ for $\tiles$ odd} \\[5mm] % *!*
\displaystyle
2\left\lceil \frac{w_\tiles}{\sqrt{2}} - \frac 1 2 \right\rceil = 2\left\lceil \sqrt{\frac{\tiles-5/2}{2}} - \frac 1 2 \right\rceil & \text{ for $\tiles$ even}.
\end{cases}
\end{equation}

@@

As for $A_{\tiles,n}$, let $\tilde A_{\tiles,n}$ denote its approximate version where the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail} is uniformly distributed on the chord rather than on the arc. This approximation entails two types of errors: the distribution of the angle $\theta$ is no longer uniform; and the shaded rectangle becomes smaller because its corner is on the chord, not on the arc. To analyze the first type of error, let $\fdp(\theta)$ denote the probability density function for the original, uniform distribution of the angle $\theta$ on the arc depicted with thick line in Figure~\ref{fig: probmax_asint_turned}. In the approximate version, the uniform distribution on the chord translates into a non-uniform distribution of $\theta$, with probability density function $\tilde\fdp(\theta)$. In the following it will be assumed that $\tiles \geq 4$. It stems from Figure~\ref{fig: probmax_asint_turned} that any point $(x,y)$ on the circular arc defined by \eqref{eq: funcirc} satisfies the equality $\cot \theta = \sqrt{2} x /(\tiles-3)$. Therefore
\begin{equation}
\label{eq: d theta d x}
\frac{\diff \theta}{\diff x} = -\frac{\sqrt{2}}{\tiles-3} \sin^2 \theta.
\end{equation}
Along this arc, with $x \in [-w_\tiles, w_\tiles]$, the minimum of $|\sin \theta|$ is attained when $x$ equals $w_\tiles$ or $-w_\tiles$. Thus, using \eqref{eq: asympt len tiles},
\begin{equation}
\label{eq: sin theta bound}
1 \geq |\sin \theta| \geq \frac{\tiles-3}{\sqrt{2}\len_\tiles} > \frac{\tiles-3}{\tiles-3/2}.
\end{equation}
The uniform distribution of the corner of the shaded rectangle on the chord is equivalent to a uniform distribution of its $x$ coordinate on the projection of the chord onto the horizontal axis. Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length $\tilde c_{\tiles,n} / |m_{\tiles,n}|$, and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
\begin{equation}
\label{eq: tilde fdp theta}
\tilde \fdp(\theta) = \frac {|m_{\tiles,n}|} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
\end{equation}
for the range of $\theta$ corresponding to the arc with thick line in Figure~\ref{fig: probmax_asint_turned}. It follows from \eqref{eq: d theta d x}--\eqref{eq: tilde fdp theta} that
% *!*
\begin{equation}
\label{eq: max min tilde pdf}
\frac {\max_\theta \tilde \fdp(\theta)} {\min_\theta \tilde \fdp(\theta)} < \left(\frac{\tiles-3/2}{\tiles-3}\right)^2.
\end{equation}
Since $\fdp(\theta)$ and $\tilde\fdp(\theta)$ are probability density functions, it cannot be the case that one of them is greater than the other for all values of $\theta$. In addition, both functions have the same support, on which $\fdp(\theta)$ is constant. Taking into account that $\tiles \geq 4$, this implies that
\begin{align}
\label{eq: tilde pdf pdf bound >}
\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &> \left(\frac{\tiles-3}{\tiles-3/2}\right)^2 = \left(1-\frac{3/2}{\tiles-3/2}\right)^2 > 1 - \frac 3 {\tiles-3/2}, \\
\label{eq: tilde pdf pdf bound <}
\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 < 1 + \frac {21/4} {\tiles-3}.
\end{align}
Thus, changing the distribution of $\theta$ from $\fdp(\theta)$ to $\tilde \fdp(\theta)$ results in $A_{\tiles,n}$ being multiplied by a factor greater than the rightmost side of \eqref{eq: tilde pdf pdf bound >} and smaller than that of \eqref{eq: tilde pdf pdf bound <}.

The second type of error in the approximation of $A_{\tiles,n}$ arises because, given the $x$ coordinate of the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail}, its $y$ coordinate is reduced from that of the arc to that of the chord. Considering the tangent line displayed in the figure (dashed line), it is easy to see that the difference between the vertical coordinate of the arc, $y$, and that of the chord, $\tilde y$, is less than the chord length, $\tilde c_{\tiles,n}$, multiplied by the maximum of the absolute value of the arc slope. The former is at most $\sqrt{2}$. The latter is upper-bounded by the value of $|\diff \funcirc(x)/\diff x|$ at $x=w_\tiles$ or $x=-w_\tiles$. From \eqref{eq: funcirc}, \eqref{eq: w} and \eqref{eq: sqrt asympt}, making use of the assumption that $\tiles \geq 4$,
\begin{equation}
\label{eq: asympt slope bound}
\begin{split}
\left| \frac{\diff \funcirc(x)}{\diff x} \right| &= \frac{|x|}{\sqrt{\len_\tiles^2-x^2}} \leq \frac{w_\tiles}{\sqrt{\len_\tiles^2-w_\tiles^2}} \leq \frac{\sqrt{2(\tiles-2)}}{\tiles-3} \\
& < \sqrt{\frac 2 {\tiles-3}} \left(1 + \frac 1 {2(\tiles-3)} \right) \leq \frac 3 2 \sqrt{\frac 2 {\tiles-3}},
\end{split}
\end{equation}
and thus
\begin{equation}
\label{eq: y: arc - chord}
0 < y - \tilde y < \frac 3 {\sqrt{\tiles-3}}.
\end{equation}
This reduction from $y$ to $\tilde y$ translates into a difference smaller than $3 / \sqrt{2(\tiles-3)}$ in each dimension of the shaded rectangle; or, since those dimensions are at most $1$, a difference in area smaller than twice that value. Considering \eqref{eq: A tiles n bound}, this introduces a (second) error in $A_{\tiles,n}$ by a factor between $1-3\sqrt{2}/\sqrt{\tiles-3}$ and $1$. Combining with the factor from the previous paragraph, $\tilde A_{\tiles,n} / A_{\tiles,n}$ is bounded for $\tiles \geq 4$ as
\begin{equation}
\label{eq: tilde A bound}
1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}} {\sqrt{\tiles-3}} < \frac{\tilde A_{\tiles,n}}{A_{\tiles,n}} < 1+\frac {21/4} {\tiles-3}.
\end{equation}

@@

%***Releer

The inequality \eqref{eq: y: arc - chord} applies, for $\tiles \geq 4$, to $y_{\tiles,n} = \funcircesc_\tiles(x_{\tiles,n}/w_\tiles)$ and $\tilde y_{\tiles,n}$. Together with \eqref{eq: P i j asympt 4 ter}, and making use of the uniform convergence of $\funcircesc_\tiles$ to $\funcircesc$, 
this implies that
\begin{equation}
\label{eq: funcircest tiles 5}
\tilde y_{\tiles,n}^3 = \funcircesc^3\left( \frac{x_{\tiles,n}} {w_\tiles} \right) + E''_{\tiles,n}
\end{equation}
where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.

From \eqref{eq: sum p t n}, \eqref{eq: P i j asympt 5} and \eqref{eq: funcircest tiles 5},
\begin{equation}
\label{eq: sum p t n aprox 1}
\tiles \probmax(\len_\tiles) = \frac{4 \sqrt{2}} {3\pi} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) + \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right),
\end{equation}
which can be rewritten as
\begin{equation}
\label{eq: sum p t n aprox 2}
\sqrt{\tiles} \probmax(\len_\tiles) = \frac{4 w_\tiles} {3\pi\sqrt{\tiles}} \frac {\sqrt{2}} {w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
+ \frac{{N_\tiles}} {\sqrt{t}} \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right).
\end{equation}
It is clear from Figure~\ref{fig: probmax_asint_turned} that
\begin{alignat}{2}
-w_\tiles &< x_{\tiles,1} && \leq -w_\tiles + \sqrt{2}, \\
w_\tiles - \sqrt{2} &\leq x_{\tiles,N_\tiles} && < w_\tiles.
\end{alignat}
Thus the values $x_{\tiles,n}/w_\tiles$, $n=1,\ldots,N_\tiles$ are equispaced with step $\sqrt{2}/w_\tiles$. In addition, $x_{\tiles,1}$ and $x_{\tiles,N_\tiles}$ tend to $-1$ and $1$  respectively as $\tiles \rightarrow \infty$. In consequence, the term
\[
\frac{\sqrt{2}}{w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
\]
in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemmann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$. Since this function is continuous the sum indeed converges to the integral \cite[section~7.2]{Abbott15}. On the other hand,
\begin{equation}
\left| \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right) \right| < e(\tiles)+e'(\tiles) + \frac{4 \sqrt{2}} {3\pi} e''(\tiles)
\end{equation}
and the right-hand side tends to $0$ as $\tiles \rightarrow \infty$. The terms $w_\tiles/\sqrt{\tiles}$ and $N_\tiles/\sqrt{\tiles}$ in \eqref{eq: sum p t n aprox 2} tend to $1$ and $\sqrt{2}$ respectively, according to \eqref{eq: w} and \eqref{eq: N tiles}. Thus, taking limits and substituting \eqref{eq: funcircesc},
\begin{equation}
\label{eq: sum p t n lim}
\lim_{\tiles \rightarrow \infty} \sqrt{\tiles} \probmax(\len_\tiles) =
\frac{4} {3\pi} \int_{-1}^1 \funcircesc^3(\zeta) \,\diff \zeta = \frac{2\sqrt{2}} {3\pi} \int_{0}^1 (1-\zeta^2)^3 \,\diff \zeta = \frac{32\sqrt{2}} {105\pi},
\end{equation}
which establishes \eqref{eq: probmax lim sqrt tiles}.

>>>>

For any $i, j \geq 2$ with $i+j-1=\tiles$, the segment has a discrete bounding rectangle with dimensions $i, j$ if and only if its second endpoint is in the tile with lower-left corner $(i-1,j-1)$ or in one of the other three symmetrical tiles with respect to the lines $x=1/2$ and $y=1/2$, i.e.~those with lower-left corners $(i-1,1-j)$, $(1-i,j-1)$ and $(1-i,1-j)$. An example is shown in Figure~\ref{fig: probmax_symmetric_tiles}. Consequently, $\probmax_{i,j}(\len)$ can be obtained by computing the probability that the second endpoint is in the tile with lower-left corner $(i-1,j-1)$ and multiplying by $4$.

@@

By Proposition~\ref{prop: len ineq i j}, the tile with lower-left corner $(i-1,j-1)$ can be visited only if
\begin{equation}
\label{eq: i-2 j-2 len}
(i-2)^2+(j-2)^2<\len^2.
\end{equation}
Thus $\probmax_{i,j}(\len)=0$ if this restriction does not hold. In the following it will be assumed that it does.

@@

Using the coordinate axes defined in Figure~\ref{fig: probmax_asint_turned}, the circular arc satisfies the equation $y = \funcirc(x)$ with
\begin{equation}
\label{eq: funcirc}
\funcirc(x) = h_\tiles - \len_\tiles + \sqrt{\len_\tiles^2-x^2},
\end{equation}
where the arc height $h_\tiles$ is 
\begin{equation}
\label{eq: h tiles}
h_\tiles = \len_\tiles - \frac{\tiles-3}{\sqrt{2}}.
\end{equation}
The $y$ coordinate of the arc for $x=x_{\tiles,n}$ will be denoted as $y_{\tiles,n} = \funcirc(x_{\tiles,n})$. The half-width of the arc, $w_\tiles$, is computed in the proof of Theorem~\ref{theo: probmax, sq} (or from the right triangle in Figure~\ref{fig: probmax_asint_turned}), and using \eqref{eq: len tiles} it can be expressed as
\begin{equation}
\label{eq: w}
w_\tiles = \sqrt{\len_\tiles^2-\frac{(\tiles-3)^2}{2}} =
\begin{cases}
\displaystyle
\sqrt{\tiles-2} & \text{ for $\tiles$ odd} \\
\displaystyle
\sqrt{\tiles-5/2} & \text{ for $\tiles$ even}.
\end{cases}
\end{equation}
The inequality
\begin{equation}
\label{eq: sqrt asympt}
\sqrt{1+\genvar} < 1 + \frac \genvar 2 \quad\text{for } s \geq -1, \ s \neq 0
\end{equation}
implies that
\begin{equation}
\label{eq: asympt w}
w_\tiles \leq \sqrt{\tiles-2} < \sqrt{\tiles} - 1/\sqrt\tiles.
\end{equation}
Similarly, \eqref{eq: len tiles} and \eqref{eq: sqrt asympt} yield, considering that $\tiles \geq 3$,
\begin{equation}
\label{eq: asympt len tiles}
\frac{\tiles-2}{\sqrt{2}} \leq \len_\tiles < \frac{\tiles-2}{\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)}
\leq \frac{\tiles-3/2}{\sqrt{2}}.
\end{equation}
The arc height $h_\tiles$ is bounded from \eqref{eq: h tiles} and \eqref{eq: asympt len tiles} as
\begin{equation}
\label{eq: h bound}
\frac {\sqrt{2}} 2 \leq h_\tiles <
% \frac 1 {\sqrt{2}} + \frac{1}{2\sqrt{2}(\tiles-2)} \leq
\frac{3\sqrt{2}}{4}.
\end{equation}
The number $N_\tiles$ of reachable tiles shown in the figure (circular markers) is easily obtained from $w_\tiles$ as
\begin{equation}
\label{eq: N tiles}
N_\tiles = 
\begin{cases}
\displaystyle
2\left\lceil \frac{w_\tiles}{\sqrt{2}} \right\rceil - 1 = 2\left\lceil \sqrt{\frac{\tiles-2}{2}} \right\rceil - 1 & \text{ for $\tiles$ odd} \\[5mm] % *!*
\displaystyle
2\left\lceil \frac{w_\tiles}{\sqrt{2}} - \frac 1 2 \right\rceil = 2\left\lceil \sqrt{\frac{\tiles-5/2}{2}} - \frac 1 2 \right\rceil & \text{ for $\tiles$ even}.
\end{cases}
\end{equation}

@@

As for $A_{\tiles,n}$, let $\tilde A_{\tiles,n}$ denote its approximate version where the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail} is uniformly distributed on the chord rather than on the arc. This approximation entails two types of errors: the distribution of the angle $\theta$ is no longer uniform; and the shaded rectangle becomes smaller because its corner is on the chord, not on the arc. To analyze the first type of error, let $\fdp(\theta)$ denote the probability density function for the original, uniform distribution of the angle $\theta$ on the arc depicted with thick line in Figure~\ref{fig: probmax_asint_turned}. In the approximate version, the uniform distribution on the chord translates into a non-uniform distribution of $\theta$, with probability density function $\tilde\fdp(\theta)$. In the following it will be assumed that $\tiles \geq 4$. It stems from Figure~\ref{fig: probmax_asint_turned} that any point $(x,y)$ on the circular arc defined by \eqref{eq: funcirc} satisfies the equality $\cot \theta = \sqrt{2} x /(\tiles-3)$. Therefore
\begin{equation}
\label{eq: d theta d x}
\frac{\diff \theta}{\diff x} = -\frac{\sqrt{2}}{\tiles-3} \sin^2 \theta.
\end{equation}
Along this arc, with $x \in [-w_\tiles, w_\tiles]$, the minimum of $|\sin \theta|$ is attained when $x$ equals $w_\tiles$ or $-w_\tiles$. Thus, using \eqref{eq: asympt len tiles},
\begin{equation}
\label{eq: sin theta bound}
1 \geq |\sin \theta| \geq \frac{\tiles-3}{\sqrt{2}\len_\tiles} > \frac{\tiles-3}{\tiles-3/2}.
\end{equation}
The uniform distribution of the corner of the shaded rectangle on the chord is equivalent to a uniform distribution of its $x$ coordinate on the projection of the chord onto the horizontal axis. Denoting the slope of the chord by $m_{\tiles,n}$, the projection has length
\[
\frac{\tilde c_{\tiles,n}} {\sqrt{1 + m_{\tiles,n}^2}},
\]
and $\tilde\fdp(\theta)$ is obtained, using the transformation theorem for continuous random variables \cite[section~5.2]{Papoulis91}, as
\begin{equation}
\label{eq: tilde fdp theta}
\tilde \fdp(\theta) = \frac {\sqrt{1 + m_{\tiles,n}^2}} {\tilde c_{\tiles,n} |\diff \theta / \diff x|}
\end{equation}
for the range of $\theta$ corresponding to the arc with thick line in Figure~\ref{fig: probmax_asint_turned}. It follows from \eqref{eq: d theta d x}--\eqref{eq: tilde fdp theta} that
% *!*
\begin{equation}
\label{eq: max min tilde pdf}
\frac {\max_\theta \tilde \fdp(\theta)} {\min_\theta \tilde \fdp(\theta)} < \left(\frac{\tiles-3/2}{\tiles-3}\right)^2.
\end{equation}
Since $\fdp(\theta)$ and $\tilde\fdp(\theta)$ are probability density functions, it cannot be the case that one of them is greater than the other for all values of $\theta$. In addition, both functions have the same support, on which $\fdp(\theta)$ is constant. Taking into account that $\tiles \geq 4$, this implies that
\begin{align}
\label{eq: tilde pdf pdf bound >}
\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &> \left(\frac{\tiles-3}{\tiles-3/2}\right)^2 = \left(1-\frac{3/2}{\tiles-3/2}\right)^2 > 1 - \frac 3 {\tiles-3/2}, \\
\label{eq: tilde pdf pdf bound <}
\frac {\tilde \fdp(\theta)} {\fdp(\theta)} &< \left(\frac{\tiles-3/2}{\tiles-3}\right)^2 = \left(1+\frac{3/2}{\tiles-3}\right)^2 \leq 1 + \frac {21/4} {\tiles-3}.
\end{align}
Thus, changing the distribution of $\theta$ from $\fdp(\theta)$ to $\tilde \fdp(\theta)$ results in $A_{\tiles,n}$ being multiplied by a factor greater than the rightmost side of \eqref{eq: tilde pdf pdf bound >} and smaller than that of \eqref{eq: tilde pdf pdf bound <}.

The second type of error in the approximation of $A_{\tiles,n}$ arises because, given the $x$ coordinate of the upper corner of the shaded rectangle in Figure~\ref{fig: probmax_asint_detail}, its $y$ coordinate is reduced from that of the arc to that of the chord. Considering the tangent line displayed in the figure (dashed line), it is easy to see that the difference between the vertical coordinate of the arc, $y$, and that of the chord, $\tilde y$, is less than the chord length, $\tilde c_{\tiles,n}$, multiplied by the maximum of the absolute value of the arc slope. The former is at most $\sqrt{2}$. The latter is upper-bounded by the value of $|\diff \funcirc(x)/\diff x|$ at $x=w_\tiles$ or $x=-w_\tiles$. From \eqref{eq: funcirc}, \eqref{eq: w} and \eqref{eq: sqrt asympt}, making use of the assumption that $\tiles \geq 4$,
\begin{equation}
\label{eq: asympt slope bound}
\begin{split}
\left| \frac{\diff \funcirc(x)}{\diff x} \right| &= \frac{|x|}{\sqrt{\len_\tiles^2-x^2}} \leq \frac{w_\tiles}{\sqrt{\len_\tiles^2-w_\tiles^2}} \leq \frac{\sqrt{2(\tiles-2)}}{\tiles-3} \\
& < \sqrt{\frac 2 {\tiles-3}} \left(1 + \frac 1 {2(\tiles-3)} \right) \leq \frac 3 2 \sqrt{\frac 2 {\tiles-3}},
\end{split}
\end{equation}
and thus
\begin{equation}
\label{eq: y: arc - chord}
0 \leq y - \tilde y < \frac 3 {\sqrt{\tiles-3}}.
\end{equation}
This reduction from $y$ to $\tilde y$ translates into a difference smaller than $3 / \sqrt{2(\tiles-3)}$ in each dimension of the shaded rectangle; or, since those dimensions are at most $1$, a difference in area smaller than twice that value. Considering \eqref{eq: A tiles n bound}, this introduces a (second) error in $A_{\tiles,n}$ by a factor between $1-3\sqrt{2}/\sqrt{\tiles-3}$ and $1$. Combining with the factor from the previous paragraph, $\tilde A_{\tiles,n} / A_{\tiles,n}$ is bounded for $\tiles \geq 4$ as
\begin{equation}
\label{eq: tilde A bound}
1 - \frac {3} {\tiles-3/2} - \frac{3\sqrt{2}} {\sqrt{\tiles-3}} < \frac{\tilde A_{\tiles,n}}{A_{\tiles,n}} < 1+\frac {21/4} {\tiles-3}.
\end{equation}

@@

The inequality \eqref{eq: y: arc - chord} applies, for $\tiles \geq 4$, to $y_{\tiles,n} = \funcircesc_\tiles(x_{\tiles,n}/w_\tiles)$ and $\tilde y_{\tiles,n}$. Together with \eqref{eq: P i j asympt 4 ter}, and making use of the uniform convergence of $\funcircesc_\tiles$ to $\funcircesc$, 
this implies that
\begin{equation}
\label{eq: funcircest tiles 5}
\tilde y_{\tiles,n}^3 = \funcircesc^3\left( \frac{x_{\tiles,n}} {w_\tiles} \right) + E''_{\tiles,n}
\end{equation}
with $|E''_{\tiles,n}| < e''(t)$, where $e''(\tiles)$ tends to $0$ as $\tiles \rightarrow \infty$.

From \eqref{eq: sum p t n}, \eqref{eq: P i j asympt 5} and \eqref{eq: funcircest tiles 5},
\begin{equation}
\label{eq: sum p t n aprox 1}
\tiles \probmax(\len_\tiles) = \frac{4 \sqrt{2}} {3\pi} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right) + \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right),
\end{equation}
which can be rewritten as
\begin{equation}
\label{eq: sum p t n aprox 2}
\sqrt{\tiles} \probmax(\len_\tiles) = \frac{4 w_\tiles} {3\pi\sqrt{\tiles}} \frac {\sqrt{2}} {w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
+ \frac{{N_\tiles}} {\sqrt{t}} \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right).
\end{equation}
It is clear from Figure~\ref{fig: probmax_asint_turned} that
\begin{alignat}{2}
-w_\tiles &< x_{\tiles,1} && \leq -w_\tiles + \sqrt{2}, \\
w_\tiles - \sqrt{2} &\leq x_{\tiles,N_\tiles} && < w_\tiles.
\end{alignat}
Thus the values $x_{\tiles,n}/w_\tiles$, $n=1,\ldots,N_\tiles$ are equispaced with step $\sqrt{2}/w_\tiles$. In addition, $x_{\tiles,1} / w_\tiles$ and $x_{\tiles,N_\tiles} / w_\tiles$ tend to $-1$ and $1$  respectively as $\tiles \rightarrow \infty$. In consequence, the term
\[
\frac{\sqrt{2}}{w_\tiles} \sum_{n=1}^{N_\tiles} \funcircesc^3 \left( \frac{x_{\tiles,n}}{w_\tiles} \right)
\]
in \eqref{eq: sum p t n aprox 2} can be interpreted as a Riemann sum that approximates the integral of $\funcircesc^3(\zeta)$ over $[-1,1]$. Since this function is continuous the sum indeed converges to the integral \cite[section~7.2]{Abbott15}. On the other hand,
\begin{equation}
\left| \frac 1 {N_\tiles} \sum_{n=1}^{N_\tiles} \left(E_{\tiles,n} + E'_{\tiles,n} + \frac{4 \sqrt{2}} {3\pi} E''_{\tiles,n} \right) \right| < e(\tiles)+e'(\tiles) + \frac{4 \sqrt{2}} {3\pi} e''(\tiles)
\end{equation}
and the right-hand side tends to $0$ as $\tiles \rightarrow \infty$. The terms $w_\tiles/\sqrt{\tiles}$ and $N_\tiles/\sqrt{\tiles}$ in \eqref{eq: sum p t n aprox 2} tend to $1$ and $\sqrt{2}$ respectively, according to \eqref{eq: w} and \eqref{eq: N tiles}. Thus, taking limits and substituting \eqref{eq: funcircesc},
\begin{equation}
\label{eq: sum p t n lim}
\lim_{\tiles \rightarrow \infty} \sqrt{\tiles} \probmax(\len_\tiles) =
\frac{4} {3\pi} \int_{-1}^1 \funcircesc^3(\zeta) \,\diff \zeta = \frac{2\sqrt{2}} {3\pi} \int_{0}^1 (1-\zeta^2)^3 \,\diff \zeta = \frac{32\sqrt{2}} {105\pi},
\end{equation}
which establishes \eqref{eq: probmax lim sqrt tiles}.
